\thispagestyle{empty}
\begin{center}
\bfseries\Large Hilbert's Axioms for Plane Geometry
\end{center}

\begin{minipage}[t]{0.46\linewidth}
\boldinline{Undefined terms}
\begin{enumerate}\itemsep2pt
  \item \emph{Points}: \ use capital letters, $A,B,C\ldots$
  \item \emph{Lines}: \ use lower case letters, $\ell,m,n,\ldots$
  \item \emph{On}: \ $A\in\ell$ is read `$A$ lies on $\ell$'
  \item \emph{Between}: \ $A*B*C$ is read `$B$ lies between $A$ and $C$'
  \item \emph{Congruence}: \ $\cong$ is a binary relation on segments or angles
\end{enumerate}

\boldinline{Axioms of Incidence}

\begin{enumerate}
	\item[I-1] For any distinct $A,B$ there exists a line $\ell$ on which lie $A,B$.
	\item[I-2] There is at most one line through distinct $A,B$ \ ($A$ and $B$ both \emph{on} the line).
\end{enumerate}\vspace{-2pt}
	
Notation: \emph{line} $\overleftrightarrow{AB}$ through $A$ and $B$\vspace{-2pt}

\begin{enumerate}\setcounter{enumi}{2}	
	\item[I-3] On every line there exist at least two distinct points. There exist at least three points not all on the same line.
\end{enumerate}


\boldinline{Axioms of Order}

\begin{enumerate}
	\item[O-1] If $A*B*C$, then $A,B,C$ are distinct points on the same line and $C*B*A$.
	\item[O-2] Given distinct $A,B$, there is at least one point $C$ such that $A*B*C$.
	\item[O-3] If $A,B,C$ are distinct points on the same line, exactly one lies between the others.
\end{enumerate}\vspace{-2pt}
	
Definitions: \emph{segment} $\cl{AB}$ and \emph{triangle} $\triangle ABC$\vspace{-2pt}

\begin{enumerate}\setcounter{enumi}{3}	
	\item[O-4] (Pasch's Axiom) Let $\triangle ABC$ be a triangle and $\ell$ a line not containing any of $A,B,C$. If $\ell$ contains a point of the segment $\cl{AB}$, then it also contains a point of either $\cl{AC}$ or $\cl{BC}$. 
\end{enumerate}

Definitions: \emph{sides} of line $\overleftrightarrow{AB}$ and \emph{ray} $\ray{AB}$
\end{minipage}
\hfill
\begin{minipage}[t]{0.48\linewidth}
\boldinline{Axioms of Congruence}

\begin{enumerate}
  \item[C-1] (Segment transference) \ Let $A,B$ be distinct and $r$ a ray based at $A'$. Then there exists a unique point $B'\in r$ for which $\cl{AB}\cong\cl{A'B'}$. Moreover $\cl{AB}\cong\cl{BA}$.
  \item[C-2] If $\cl{AB}\cong\cl{EF}$ and $\cl{CD}\cong\cl{EF}$, then $\cl{AB}\cong\cl{CD}$.
  \item[C-3] If $A*B*C$, \ $A'*B'*C'$, \ $\cl{AB}\cong\cl{A'B'}$ and $\cl{BC}\cong\cl{B'C'}$, then $\cl{AC}\cong\cl{A'C'}$.
  \end{enumerate}
	
Definitions: \emph{angle} $\angle ABC$\vspace{-2pt}

\begin{enumerate}\setcounter{enumi}{3}
  \item[C-4] (Angle transference) \ Given $\angle BAC$ and $\ray{A'B'}$, there exists a unique ray $\ray{A'C'}$ on a given side of $\overleftrightarrow{A'B'}$ for which $\angle BAC\cong\angle B'A'C'$.
  \item[C-5] If $\angle ABC\cong\angle GHI$ and $\angle DEF\cong\angle GHI$, then $\angle ABC\cong\angle DEF$. Moreover, $\angle ABC\cong\angle CBA$.
  \item[C-6] (Side-angle-side) Given triangles $\triangle ABC$ and $\triangle A'B'C'$, if $\cl{AB}\cong\cl{A'B'}$, $\cl{AC}\cong\cl{A'C'}$, and $\angle BAC\cong\angle B'A'C'$, then the triangles are congruent.\footnotemark{}
\end{enumerate}

\boldinline{Axiom of Continuity}\phantom{bob}\smallbreak

Suppose $\ell$ is partitioned into non-empty subsets $\Sigma_1,\Sigma_2$ such that no point of $\Sigma_1$ lies between two points of $\Sigma_2$ and vice versa.\par
Then there exists a unique point $\mathcal O\in\ell$ satisfying $P_1*\mathcal O*P_2$, if and only if $\mathcal O\neq P_1$, $\mathcal O\neq P_2$ and one of $P_1$ or $P_2$ lies in $\Sigma_1$ and the other in $\Sigma_2$.\medbreak



\boldinline{Playfair's Axiom}\phantom{bob}\smallbreak

Definition: \emph{parallel lines}\smallbreak

Given a line $\ell$ and a point $P$ not on $\ell$, there exists at most one line through $P$ parallel to $\ell$.
\end{minipage}

\footnotetext{Its sides/angles are congruent in pairs. We extend congruence to other geometric objects similarly.}

\clearpage



